---
title: "**Analysis of Titanic shipwreck**"
author: "**Mohamed Faadil**"
output:
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(digits = 4 , scipen = 4)
setwd("/Users/faadil/Desktop/Boston_University/CS-555/Project/Project_555") 
library(tidyverse)
#install.packages('caTools')
library(caTools)
#install.packages("ggcorrplot")
library("ggcorrplot")
#install.packages("fastDummies") 
library(fastDummies)
#install.packages("lsmeans")
library("lsmeans")
#install.packages("ggfortify") ## installation required
library(ggfortify)
#install.packages("olsrr") ## installation required
library("olsrr")
library(car)
library(GGally)
library(PerformanceAnalytics)

# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
knitr::opts_chunk$set(opts.label="kill_prefix")
```

```{css, echo=FALSE}
.main-container {
    max-width: 100%;
}
pre {
  font-size: 20px;
  font-weight: bold;
  font-family: Arial;
}
body, td {
  font-size: 20px;
  font-family: Arial;
}
h1,h2,h3,h4,h5,h6 {
  font-family: Arial;
}
```
## **Data Set Description**


This Data set has 891 rows and 12 columns. There are 11 predictor variables and 1 target variable.

```{r data set description, echo=FALSE}
#import data 
df <- read.csv("/Users/faadil/Desktop/Boston_University/CS-555/Project/Project_555/titanic_population.csv", header = TRUE)
head(df)
attach(df)
df_clean <- (df[c(2,3,5,6,7,8,10,12)])
df_clean <- na.omit(df_clean)
df_clean <- head(df_clean,500)
```

## **Correlation Plot**
```{r correlation plot, fig.align='center', fig.width=20, fig.height=13, echo=FALSE, results='hide'}
chart.Correlation(df[,c(1,2,3,6,7,8,10)], histogram = TRUE, method = "pearson")
```

## **Explanatory Data Analysis**



``` {r Explaratory Data Analysis}

df_dummy <- fastDummies::dummy_cols(df_clean,
select_columns = c("Sex","Embarked"))
colnames(df_dummy)
df_dummy <- df_dummy[c(1,2,4,5,6,7,9,10,12,13,14)]
corr <- (cor(df_dummy[,-9]))
ggcorrplot(corr, hc.order = F, outline.col = "white",
           lab = TRUE, title= "Correlation heatmap")

summary(df_clean)
summary(subset(df_clean, df_clean$Survived ==0))
summary(subset(df_clean, df_clean$Survived ==1))

#Bar Plots Survived vs Non-Survived
ggplot(df_clean, aes(x=Survived, y=Age, fill=Sex)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=2, notch=FALSE) +
  labs(title="Box plot",
       x = "Survived",
       y = "Age")

ggplot(df_clean, aes(x=Survived, y=Age, fill=Sex)) + 
  geom_bar(stat="identity") +
  labs(title="Bar plot",
       x = "Survived",
       y = "")
ggplot(df_clean, aes(x=Pclass, y=Survived, fill=Sex)) + 
  geom_bar(stat="identity") +
  labs(title="Bar plot",
       x = "Passenger Class",
       y = "")
```


## **Logistic Regression**

Building logistic regression classification models with below classes.

logistic regression of survived passengers as a function of passenger class. Passenger class is significant in determining the survival at 95% confidence interval

``` {r Logistic Regression}

## Passenger Class:
m <- glm(Survived~Pclass, 
         data = df_dummy , family = binomial)
summary(m)

```



##Adjusting for age:

Ran anova test on the following logistic regression model.

Categorical variable: passenger class

Covariate : Age

They are both significant in determining the survival of passenger

``` {r Logistic RRegression}
library(car)
Anova(glm(Survived~Pclass + Age, 
         data = df_dummy , family = binomial), type = 3, test.statistic = "F")
```


## Regression with Gender

Looking below we can determine gender is significant in determining the survival of passenger

``` {r Logisticc Regression}

m <- glm(Survived~Sex_female, 
         data = df_dummy , family = binomial)
summary(m)

```


#Adjusting for age:

Ran anova test on the followinglogistic regression model.

Categorical variable: Sex

Covariate : Age

Both are significant together in determining survival

``` {r Annova Sex}
Anova(glm(Survived~Sex_female+Age, 
          data = df_dummy , family = binomial), type = 3, test.statistic = "F")

```


## Family

Looking below we can determine Family is significant in determining the survival of passenger

``` {r Anova Family}
m <- glm(Survived~Parch, 
         data = df_dummy , family = binomial)
summary(m)
```


#Adjusting for age:

Ran anova test on the following logistic regression model.

Categorical variable: family

Covariate : Age

Both are not significant at 95 % CI to determine survival

``` {r Anova Family adjusting age}
Anova(glm(Survived~Parch+Age, 
          data = df_dummy , family = binomial), type = 3, test.statistic = "F")

```


## Siblings

Looking below we can determine sibling is not significant 

``` {r Anova with Siblings}

m <- glm(Survived~SibSp, 
         data = df_dummy , family = binomial)
summary(m)

```


#Adjusting for age:

Ran anova test on the following logistic regression model.

Categorical variable: Sibling

Covariate : Age

Both are not significant at 95 % CI to determine survival

``` {r Anova sibling with age}
Anova(glm(Survived~SibSp+Age, 
          data = df_dummy , family = binomial), type = 3, test.statistic = "F")

```

## **Logistic Regression Classification Model**

Fare is not significant in determining survival at  95% CI.

``` {r Final Logistic Regression}
#building the final logistic regression classification model

set.seed(64)
split = sample.split(df_dummy$Survived, SplitRatio = 0.8)
training_set = subset(df_dummy, split == TRUE)
test_set = subset(df_dummy, split == FALSE)
training_set[-1] = scale(training_set[-1])
test_set[-1] = scale(test_set[-1])

classifier = glm(formula = Survived ~Pclass+Sex_female+Fare+Age,
                 family = binomial,
                 data = training_set)

summary(classifier)
```

## **Predicting Test Set Results**

We have used threshold as 0.5 to perform classification.

``` {r Test Reuslts}
# Predicting the Test set results
model.probs <- predict(classifier, test_set, type = "response")
model.pred <- rep(0, length(model.probs))
model.pred[model.probs > 0.5] <- 1
```

## **Confusion Matrix**

TN = 48
FP = 10
FN = 11
TP = 31

```{r Confusion Matrix}
# Making the Confusion Matrix
table(model.pred, test_set$Survived)

# Accuracy 
1- mean(model.pred != test_set$Survived)

# ROC curve
#install.packages("pROC")
library(pROC)
g <- roc(test_set$Survived ~ model.probs)
print(g)
plot(g, main = "ROC curve") 



```

## **Summary**

We conclude that Passenger Class, gender were the significant parameters in survival. Initially, the female gender parameter showed some sort of correlation with passenger survival. These parameters were: Pclass, gender, age, Fare and # of family members. ANCOVA was performed, adjusting for age on such parameters, and finally it was found that age wasn’t a significant predictor if gender is the predictor. And, finally, Pclass, gender, and fare were the only significant predictors, once adjusted for age.
Based on a logistic regression model, using Pclass, gender, Fare as its features, we could predict the survival with an accuracy of 79%. Another assessment of this model, the ROC curve, showed good results as the area under the ROC curve was 0.815.

